### 2017-11-22 paper.tex
Update on Overleaf.

\documentclass{llncs}
\usepackage{epsfig}
\usepackage{graphicx}

\newtheorem{mydef}{Def.}
\newtheorem{myhyp}{Hypothesis}

\usepackage{url}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{hhline}
\usepackage{caption}
\usepackage{subcaption}

%\pagestyle{empty}
\pagestyle{headings}

\begin{document}

%----------------------------------------------------------------------

\title{Using Metric Space Indexing for Complete and Efficient Record Linkage}

%----------------------------------------------------------------------

\author{Submitted for double-blind review}

%  Hidden For Blind Review
\iffalse

\author{{\"O}zg{\"u}r Akg{\"u}n\inst{1} \and Alan Dearle\inst{1} \and\\
Graham Kirby\inst{1} \and Peter Christen\inst{2}}

\institute{School of Computer Science, University of St Andrews,\\
St Andrews, Scotland. Contact: \email{ozgur.akgun@st-andrews.ac.uk}
\and
Research School of Computer Science, The Australian National University,\\
Canberra, Australia. Contact: \email{peter.christen@anu.edu.au}}

\fi
%----------------------------------------------------------------------

\maketitle

\begin{abstract}

Record linkage is the process of identifying records that refer to the
same real-world entities, in situations where entity identifiers are
unavailable. Records are linked on the basis of similarity between
common attributes, with every pair being classified as a link or
non-link depending on their degree of similarity. Record linkage is
usually performed in a three-step process: first groups of similar
candidate records are identified using indexing, pairs within the same
group are then compared in more detail, and finally classified.
%
Even state-of-the-art indexing techniques, such as Locality Sensitive
Hashing, have potential drawbacks. They may fail to group together some
true matching records with high similarity. Conversely, they may group
records with low similarity, leading to high computational overhead.
%
We propose using metric space indexing to perform \emph{complete} record
linkage, which results in a parameter-free record linkage process
combining indexing, comparison and classification into a single step
delivering complete and efficient record linkage. Our experimental
evaluation on real-world datasets from several domains shows that
linkage using metric space indexing can yield better quality than
current indexing techniques, with similar execution cost, without the
need for domain knowledge or trial and error to configure the process.

\end{abstract}

\keywords Entity resolution; data matching; similarity search;
         blocking.

%----------------------------------------------------------------------

\section{Introduction}
\label{sec-intro}

Record linkage, also known as entity resolution, data matching and
duplicate detection~\cite{Chr12}, is the process of identifying and
matching records that refer to the same real-world entities within or
across datasets. The entities to be linked are often people (such as
patients in hospital or customers in business datasets), but record
linkage can also be applied to link consumer products or bibliographic
records~\cite{Chr12}. Record linkage is commonly challenged by the lack
of unique entity identifiers (keys) in the datasets to be linked, which
prevents the use of a database join. Instead, the linkage of records
requires the comparison of the common attributes (or fields) that are
available within the datasets. For datasets that contain information
about individuals, these attributes include the names, addresses,
dates of birth, and so on, of individuals.

To overcome data quality issues such as typographical errors and
variations (which are common in name and address values~\cite{Chr12}),
approximate string comparison functions (such as edit distance, the
Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used to
compare pairs of records, leading to a vector of similarities (one
similarity per attribute compared) for each pair. These similarity
vectors are then used to classify the compared record pairs into
\emph{links} (where it is assumed both records in a pair correspond to
the same real-world entity) and \emph{non-links} (where the records are
assumed to correspond to two different entities). Various
classification methods have been employed in record
linkage~\cite{Chr12,Don15}, ranging from simple threshold-based to
sophisticated clustering, supervised classification techniques, and
active learning approaches~\cite{Wan15}.

Besides the issues of a lack of unique entity identifiers and data
quality (which will affect linkage quality), record linkage is also
challenged by the increasing sizes of the datasets to be
linked~\cite{Don15}. To avoid full pair-wise comparison of all possible
record pairs (quadratic in the sizes of the datasets to be linked),
blocking techniques, commonly known as \emph{indexing}~\cite{Chr12b},
are used. These split the datasets into smaller blocks in a
computationally efficient way, such that records that are likely to
correspond to the same entity are grouped into the same block. Only
records within the same block are then compared in more detail.

While indexing techniques facilitate efficient linkage of very large
datasets~\cite{Don15}, they generally achieve scalability at the cost
of reduced linkage quality, because potentially true matching record
pairs are removed in the indexing step, leading to a reduction in
recall~\cite{Chr12}. A variety of indexing
techniques, discussed in more detail in the following section, have
been proposed, ranging from simple phonetic based blocking~\cite{Chr12}
and sorting of the datasets~\cite{Dra12} to locality sensitive hashing
based techniques~\cite{Kim10,Steorts2014}. Techniques for
unsupervised~\cite{Kej13,Ram15} and supervised~\cite{Bil06,Mic06}
learning of optimal blocking schemes have also been proposed.

%----------------------------------------------------------------------

\begin{figure}[!t]
  \centering
  \includegraphics[width=1.0\textwidth]{figures/linkage-process}
  \caption{Overview of the steps of the traditional record linkage
           process (left side) and our proposed metric space
           indexing based approach (right side), as described in
           Sect.~\ref{sec-intro}, where records from two datasets,
           $\mathbf{D}_A$ and $\mathbf{D}_B$, are being linked.}
           \label{fig-rl-process}
\end{figure}

%----------------------------------------------------------------------

Traditional record linkage systems that perform indexing prior to
comparison and classification, as illustrated on the left in
\Cref{fig-rl-process}, add a further practical complexity to the
process. Indexing, comparison and classification are often conducted
using algorithms and parameters selected based on technical and
domain expertise of the user of the record linkage system, followed
by a manual assessment (auditing) of the linkage
outcomes~\cite{Chr12}. If the quality of the resulting links is not
good enough for a certain application, the linkage process needs to
be repeated with different parameter settings and possibly also
alternative algorithms. This can result in a time-consuming and
labour-intensive iterative process. A major challenge is often the
heuristic nature of indexing, where the choice of a certain indexing
technique and its parameters (including which attributes to use in indexing) will determine the final outcome of a linkage.

This paper focuses on approaches that use a similarity threshold to
classify links. Such techniques are fundamentally limited by the extent
to which true matching records are similar, and true non-matches are
dissimilar---this is dataset-dependent. Within this domain, we define
a technique to be \emph{complete} if it is guaranteed to identify all
record pairs within the specified similarity threshold. Commonly used
indexing techniques are incomplete, since they reduce computational
cost at the expense of potentially overlooking some true matches. By
definition, incomplete techniques yield lower recall than complete
techniques. Conversely, and perhaps counter-intuitively, complete
techniques can yield lower precision with some datasets. This is
discussed further in Sect.~\ref{sec-approach}.

\emph{Metric space indexing} (MSI) is a complete technique with much
lower computational overheads than a brute force approach. It allows
indexing, comparison and classification to be combined into a single
step, as illustrated in the right-hand side in
Fig.~\ref{fig-rl-process}, making the overall record linkage process
simpler, more efficient and more effective than incomplete indexing
approaches.

\smallskip
\textbf{Contribution:} The primary contribution of this paper is the
novel application of MSI to achieve complete and efficient record
linkage, without the need for complex parameter tuning. We evaluate our
approach on several real-world datasets from diverse domains and
demonstrate its advantages over existing indexing techniques for
record linkage.

%----------------------------------------------------------------------

% For final paper:
% The motivation for this work is the Digitising Scotland project
% (Dibben 2012), which is in the process of transcribing and linking all
% the vital events recorded in Scotland between 1856 and 1977. This data
% set will, when complete, include around 14 million birth records, 11
% million death records and 4 million marriage records. As part of the
% work, certain data fields (locations, occupations and causes of death)
% are also being classified to the relevant standard coding schemes.

%----------------------------------------------------------------------

\section{Related Work}
\label{sec-related}

We now briefly review relevant work in the areas of indexing for record
linkage (for recent surveys see~\cite{Chr12b,Pap16}), and metric space
indexing~\cite{Zezula2010}.

Techniques to link records across datasets have been investigated for
over five decades~\cite{Fel69,New59}, with the scalability of linking
being an ongoing challenge as datasets grow in size and complexity.
Traditional blocking~\cite{Chr12b} uses a set of attributes (known as a
\emph{blocking key}) to insert records that share the same value(s) in
their blocking key into the same block. Only records within the same
block are  compared to each other. To overcome variations and
misspellings, the values used in blocking keys can be phonetically
encoded using functions such as Soundex, NYSIIS, or
Double-Metaphone~\cite{Chr12}. These convert a string into a code
according to its pronunciation, assigning the same code to similar
sounding names (such as `Gail' and `Gayle'). Multiple blocking keys
may also be used to deal with the problem of missing attribute values.

A different approach to indexing is the sorted neighbourhood
method~\cite{Mon96}, where the datasets to be linked are sorted
according to a \emph{sorting key} (usually a concatenation of the values
from several attributes). A sliding window is then moved over the
datasets and only records within the window are compared. Techniques
that adaptively shrink or expand the window size based on the
characteristics of the sorting key values have been shown to improve
both linkage efficiency and quality~\cite{Dra12} over
%other sliding window
approaches that use fixed size windows.

These blocking techniques are heuristics, commonly requiring domain
knowledge, such as the choice of appropriate blocking or sorting keys.
Poor choices of blocking attributes result in records being inserted
into inappropriate blocks, and thus true matches being missed. As a
result, such techniques can lead to \emph{incomplete} linkage.
Conversely, many of the record pairs compared in a block may turn out
to have low similarity, corresponding to non-matches, resulting in
\emph{inefficient} linkage.

Locality sensitive hashing (LSH), originally proposed to allow efficient
nearest-neighbour search in high-dimensional spaces~\cite{Ind98}, has
been employed in record linkage as an indexing technique where attribute
values are hashed multiple times, and blocks are created from those
records that share some hash values. \emph{HARRA}~\cite{Kim10} is a
record linkage approach based on MinHash~\cite{Broder1997} and LSH
which blocks, compares, and then merges linked records in an iterative
fashion, where merged records are re-hashed to improve overall linkage
quality. \cite{Steorts2014} evaluates two LSH variations, concluding
that in order to get good results, LSH methods must be tuned to the
particular datasets being linked. This requires good quality ground
truth data which may be unavailable or expensive to obtain.

Metric space indexing (MSI) techniques~\cite{Zezula2010} provide
indexing structures to support the comparison of records in one set
with those in another, typically also offering similarity search
operations. To create an MSI data structure, it is necessary to define
a distance measure between records, with certain properties including
the \emph{triangle inequality}~\cite{Zezula2010}. Similarity search
operations include \emph{range-search(\textbf{q},d)}, where all
records $\mathbf{r}$ in a database $\mathbf{D}$ within a distance $d$
of a query record $\mathbf{q}$ are identified;
\emph{nearest-neighbour(\textbf{q})}, returning the record $\mathbf{r}
\in \mathbf{D}$ with smallest distance to $\mathbf{q}$; and
\emph{nearest-n(\textbf{q},n)}, returning the $n$ closest records in
$\mathbf{D}$ to $\mathbf{q}$.
%\cite{Zezula2010} gives an overview of MSI techniques.
Here we choose one metric space indexing structure, the
M-tree~\cite{Ciaccia97indexingmetric}, and investigate its efficacy
for record linkage.

The M-tree is a dynamically balanced tree structure. Every node contains
a reference to a record being indexed, a pointer to its parent, the
distance to its parent, and the node's radius. The radius of a node is
the distance from it to its furthest child. For a parent node with
radius \textit{r}, all its children may be visualised as being contained
within a ball of radius \textit{r} from it.

A threshold-based linkage method using R-trees~\cite{Hjaltason1998}
was described in~\cite{Li2006}. It demonstrates that high linkage
quality can be achieved using Jaccard similarity over selected record
attributes. \cite{Ciaccia97indexingmetric} shows that M-trees are
almost always more efficient than R-trees, hence their use in the
experiments described in this paper.

%----------------------------------------------------------------------

\section{Approach}
\label{sec-approach}

We address the following general record linkage problem: for two
datasets $\mathbf{D}_A$ and $\mathbf{D}_B$, we wish to find, for each
record in $\mathbf{D}_A$, all the records in $\mathbf{D}_B$ that match
it with regard to a certain distance threshold, $d$ (i.e.\ have a
distance of $d$ or less). We compare several algorithms for record
linkage: traditional blocking, one incomplete similarity search method,
LSH-MinHash, and one complete method, M-tree. We also use a
simple complete brute force technique as a baseline for comparison,
although this can only feasibly be applied to the smallest of our
datasets. All experiments have a number of parameters to configure
the search space and algorithm behaviour, including a
\emph{distance threshold}, $d$, specifying a maximum distance,
equivalent to a minimum similarity, for two records to be classified
as a link (i.e.\ referring to the same entity).

\smallskip
\textbf{Brute force:}
Two nested loops are used to compare every record in
$\mathbf{D}_A$ with every record in $\mathbf{D}_B$. Each pair is
classified as a link if the distance between the records is equal to
or less than the threshold $d$. This approach is guaranteed to identify
all links, with complexity $ O(|\mathbf{D}_A| \times |\mathbf{D}_B|)$.

\smallskip
\textbf{Traditional blocking:}
The parameters of traditional blocking are the set of blocking keys
and the (optionally) phonetic
% Peter: normalisation?
encodings applied to each attribute. These are selected as described
in~\cite{Chr12b}, exploiting knowledge of the domain and of the
records being linked, and chosen with the intention of giving the
best possible results for the datasets. Each record in $\mathbf{D}_A$
is placed into the appropriate block based on its blocking key value.
The algorithm then iterates over the records in $\mathbf{D}_B$, and
for each one, compares it with each of the records from $\mathbf{D}_A$
in the block with the same blocking key value.

\smallskip
\textbf{LSH-MinHash:}
The parameters for LSH-Minhash are~\cite{Broder1997} \emph{shingle
size} ($l_{ss}$), \emph{band size} ($l_{bs}$) and \emph{number of
bands} ($l_{nb}$). Each record in $\mathbf{D}_A$ is placed in an
\emph{LSH-MinHash} data structure. First, all the attributes of the
record are concatenated, and the result \emph{shingled} into a set of
n-grams with $n = l_{ss}$. Next, a set of deterministically generated
hash functions are applied to each n-gram in the set and the smallest
result (the MinHash) of each hash application is added to a signature
for the record. The number of hashes used, and thus the size of the
signature, is set to $l_{nb} \times l_{bs}$. Finally, the signature is
split into $l_{nb}$ bands and the values from each band are hashed
again to create a number of keys. The original record is added to a
map associated with each of the keys.

To perform linkage, the algorithm iterates over the records in
$\mathbf{D}_B$. Each record is hashed using the procedure described
above, to obtain a set of keys. For each of these keys, the key is
looked up in the LSH-MinHash data structure, and the records from
$\mathbf{D}_A$ associated with it added to the result set. Finally,
the record from $\mathbf{D}_B$ is compared in turn with each record
in the result set using the same distance function and threshold, $d$,
as with the other approaches, with the pair being classified as a link
or non-link based on their distance.

In some circumstances, incomplete approaches such as traditional
blocking and LSH-MinHash can yield higher precision than complete
techniques. This can occur when the datasets contain a significant
number of non-matches that nonetheless have high similarity. In this
situation, the fact that an incomplete technique omits consideration
of some potential links can serve to improve precision, since a
classification decision based on a certain similarity threshold is
incorrect for high-similarity non-matches. By definition, recall can
never be higher for incomplete techniques.

\smallskip
\textbf{M-tree:} Besides the distance threshold, $d$, the M-tree
linkage algorithm has no additional parameters. In a similar manner
to the \emph{LSH-MinHash} approach, each record in $\mathbf{D}_A$ is inserted into an M-tree. To perform linkage, the algorithm iterates
over each record $\mathbf{b} \in \mathbf{D}_B$. A
\emph{range-search(\textbf{b},d)} operation is performed on the
M-tree, passing the distance threshold $d$ as the second parameter.
All the returned records are directly classified as links.

%----------------------------------------------------------------------

\section{Experiments and Results}
\label{sec-exp}

We now describe the datasets we use in our evaluation, the set-up
employed to evaluate our proposed metric indexing approach and compare
it with traditional blocking techniques, and we then present and discuss
our results~\footnote{The raw data, additional figures and the source
code required to run these experiments can be downloaded from:
(link removed for double blind review).}.

%----------------------------------------------------------------------

\begin{table}[t]
\caption{Characteristics of datasets used in the experiments.}
 \label{table-datasets}
  \centering
  \begin{scriptsize}
  %\addtolength{\tabcolsep}{-0.5pt}
  \begin{tabular}{ccccc}
  \hline\noalign{\smallskip}
  Dataset~ & ~Records in~& ~Records in~ & ~Number of true~& ~Entities \\
    name(s)  & ~dataset $\mathbf{D}_A$~ & ~dataset $\mathbf{D}_B$~ &
    matching pairs & linked \\
  \noalign{\smallskip} \hline \noalign{\smallskip}
  Cora & 1,295 & 1,295 & 17,184 & Publication--Publication\\
  ~Isle of Skye~ & 17,612 & 12,284& 2,900 & Birth--Death \\
  Kilmarnock & ~38,430~ & ~23,714~ & ~8,300~ & ~Birth--Death~ \\
  \noalign{\smallskip} \hline
  \end{tabular}
  \end{scriptsize}
\end{table}

%----------------------------------------------------------------------

%\smallskip
%\textbf{Datasets:}
\label{sec-data}
We used three datasets from two domains in our experiments, as
summarised in Table~\ref{table-datasets}. The first is
\emph{Cora}~\cite{Cora2017}, which contains 1,295 records that refer
to 112 machine learning publications. Cora is commonly used as a
benchmark dataset in the literature for assessing linkage algorithms.
Ground truth is provided via a unique \emph{paper\_id} identifier of
the form ``blum1993". In this experiment linkage is performed over
the same set of records (i.e.\ a de-duplication~\cite{Chr12}).

The other two datasets are historical Scottish records of vital events
(birth, marriages and deaths) one registered on the \emph{Isle of
Skye}, a rural district, and the other records from
\emph{Kilmarnock}, an industrial town. These datasets were created by
historical demographers who extensively curated and linked both
datasets~\cite{reid2006,reid2002}. Both datasets include the names,
gender, addresses of individuals and their parents. Ground truth was
generated by the demographers based on their extensive domain
knowledge.

In all of our experiments we use a single distance metric which is the
sum of the Levenshtein~\cite{Levenshtein66} edit distances between the
set of attributes being compared.

\subsection{Cora Results}

In this section, we perform linkage on the Cora dataset using all four
approaches presented in this paper: brute force, traditional blocking,
LSH and M-tree, using several selected configurations for the blocking
and LSH approaches. The distance threshold is varied between $0 \le d
\le 250$~\footnote{Relatively high Levenshtein edit distances are
included since Cora includes a number of low-similarity true
matches.}. For traditional blocking, the following Cora attributes are
used individually as blocking keys: \emph{author}, \emph{title},
\emph{venue}, \emph{location}, \emph{publisher} and \emph{year}. We
also use a combined blocking key comprising all attributes.

\Cref{cora-quality} shows the precision, recall, and F-measure~\cite{Chr12} for various thresholds. As expected, low 
thresholds give high precision and low recall, and the reverse for high
threshold values. Brute force and M-tree give identical results, as
expected. The best linkage quality, with an F-measure of around 0.7,
is achieved by a number of linkers, including brute force, M-tree, blocking on \emph{authors} and blocking on all attributes, and two
of the LSH configurations. All of these give similar overall results,
apart from blocking on \emph{authors}, which gives much better quality
at very high distance thresholds. This is due to the incomplete
nature of the approach, avoiding comparisons of significant numbers
of high-similarity non-matches and thus avoiding these becoming false
positives and keeping precision high.

For a more detailed investigation of selected linkers, the brute force
approach is used to establish a good threshold value for the Cora
dataset. The maximum F-measure is observed at a threshold value of
$d=70$. This value is dataset-dependent; for different datasets the
maximum F-measure will occur at different thresholds. In the rest of
this section we fix the threshold value at $d=70$.

\Cref{comparison-of-results-cora} shows greater detail for selected
linkers, showing the parameters for the experiment, the number of
distance comparisons made, and the precision, recall and F-measure
achieved by each algorithm. In the \emph{Linker} column the algorithm
name is followed by its parameters: for LSH the number of the bands
followed by the band size, and for traditional blocking the attributes
used for blocking. The number of distance comparisons is reported as a
machine-independent proxy for execution cost, since code profiling
shows that distance calculations are dominant.

M-tree yields the same linkage quality as brute force, although using a
significantly lower number of comparisons. This is as expected, since
both techniques are complete. Several of the incomplete linkers give
similar quality, for example \emph{LSH-2-2}, \emph{LSH-5-2},
\emph{LSH-10-2} and \emph{Block-combined}. These, and a number of other
incomplete linkers, give better precision than the complete techniques.
This is due to high-similarity non-matches, as discussed in
Sect.~\ref{sec-approach}. Although several of the incomplete linkers
give as good quality as M-tree, and in some cases at lower cost, this is
offset by the need to select appropriate configuration parameters. Some
other linkers give very poor results.

%----------------------------------------------------------------------

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
\vspace{-6mm}
\caption{Brute Force}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
\vspace{-6mm}
\caption{M-tree}
\end{subfigure} \vspace{5mm}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
\vspace{-6mm}
\caption{Blocking on `authors' attribute}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
\vspace{-6mm}
\caption{Blocking on `title' attribute}
\end{subfigure} \vspace{5mm}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
\vspace{-6mm}
\caption{Blocking on `publication year' attr.}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
\vspace{-6mm}
\caption{Blocking using the union of all attrs.}
\end{subfigure} \vspace{5mm}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-2}
\vspace{-6mm}
\caption{LSH using 2 bands of size 2}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-10}
\vspace{-6mm}
\caption{LSH using 2 bands of size 10}
\end{subfigure} \vspace{5mm}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-2}
\vspace{-6mm}
\caption{LSH using 10 bands of size 2}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-10}
\vspace{-6mm}
\caption{LSH using 10 bands of size 10}
\end{subfigure}

\caption{Linkage results on the Cora dataset.}
\label{cora-quality}
\end{figure}

%----------------------------------------------------------------------

\begin{table}[ht]
\caption{Linkage quality on Cora dataset with distance threshold
  $d= 70$.}
  \label{comparison-of-results-cora}
  \centering
  \begin{footnotesize}
  \begin{tabular}{lrrrr} \hline\noalign{\smallskip}
  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
%\parbox{0.2\textwidth}{\centering Linker} &
%\parbox{0.2\textwidth}{\centering ~~~Comparisons} &
%\parbox{0.17\textwidth}{\centering ~~~Precision} &
%\parbox{0.17\textwidth}{\centering ~~~Recall} &
%\parbox{0.17\textwidth}{\centering ~~~F1 Measure} \\
  \noalign{\smallskip} \hline \noalign{\smallskip}
Brute Force        & 1,677,025                 & 0.84      & 0.57   & 0.68  \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree              &  902,693                  & 0.84      & 0.57   & 0.68       \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2            &  192,199	& 0.95	& 0.47	& 0.63       \\
LSH-5-2            &  342,849	& 0.91	& 0.55	& 0.69       \\
LSH-10-2           &  513,947	& 0.88	& 0.57	& 0.69       \\
LSH-2-5            &  14,329	& 0.99	& 0.28	& 0.43      \\
LSH-5-5            &  22,057	& 0.99	& 0.36	& 0.53       \\
LSH-10-5           &  26,167	& 0.98	& 0.4	& 0.57      \\
LSH-2-10           &  4,711	& 1	& 0.15	& 0.27    \\
LSH-5-10           &  6,501	& 1	& 0.19	& 0.32      \\
LSH-10-10          & 10,627  	& 0.99	& 0.27& 	0.43       \\
\noalign{\smallskip} \hline \noalign{\smallskip}
Block-year         &  115,893                  & 0.99      & 0.35   & 0.51       \\
Block-authors      &   11,039                   & 0.94      & 0.16   & 0.28       \\
Block-title        &   27,407                   & 0.95      & 0.42   & 0.58       \\
Block-venue        &   36,647                   & 0.85      & 0.29   & 0.44       \\
Block-location     & 1,009,957                 & 0.83      & 0.43   & 0.57       \\
Block-publisher    &  833,079                  & 0.85      & 0.44   & 0.58       \\
Block-combined     & 1,214,269                 & 0.84      & 0.56   & 0.67       \\
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

%----------------------------------------------------------------------

\subsection{Demographic Dataset Results}

Birth records were linked to death records, separately for the Skye and
Kilmarnock datasets, using M-tree and a range of LSH configurations. It
was not computationally feasible to run the brute force linker. Of the
incomplete linkers, LSH was selected as it gave slightly better results
for Cora. The shingle size was set to $l_{ss}=2$ for all the LSH
experiments reported, as this was found to give good results and LSH
was not especially sensitive to this parameter. Results for other
shingle sizes are omitted from this paper for brevity. A lower range
of distance thresholds was explored, based on domain knowledge of the
datasets.

\Cref{demography-quality} plots (a) and (b) show the M-tree precision,
recall, and F-measure for various thresholds. In both datasets, the best
F-measure values are obtained with a low distance threshold of $d=2$.
Plots (c) and (d) compare the F-measure curves for M-tree with those
obtained from a range of LSH configurations. The best F-measure value
for M-tree is higher than that of any of the LSH configurations, for
both datasets. This demonstrates both the competitiveness of M-tree
with respect to linkage quality, and its important characteristic of
being parameter-free---the linkage quality is obtained without the
need to tune for the dataset.

Tables~\ref{comparison-of-results-demography-skye}
and~\ref{comparison-of-results-demography-kili} show greater detail for
selected linkers. In both cases the F-measure achieved is better for
M-tree than any of the LSH linkers. The better linkage quality achieved
by M-tree is largely due to recall for M-tree being much higher than for
any of the LSH configurations. In most cases, LSH out-performs M-tree in
terms of precision. More significantly, LSH linkage quality is heavily
dependent on the configuration parameters. For plausible settings for
the number of bands and band size, F-measure varies from $0.01$
(extremely poor) to $0.47$ (relatively good) for Skye and from $0.03$
to $0.49$ for Kilmarnock. In both cases \emph{LSH-10-2} performs best,
but since this is data-dependent there is no guarantee that these
parameters would work well with another dataset.

The number of distance comparisons varies dramatically among the various
linkers. M-tree always performs the most comparisons, since they are
intrinsic to the \emph{range-search} algorithm. The core part of the LSH
linker performs Jaccard similarity comparisons and hashing; distance
comparisons are only performed in the final step to determine whether a
candidate pair is a link. The LSH configurations yielding the best
results perform distance comparisons of the same order of magnitude as
M-tree. This indicates that the good LSH linkers return many candidates
beyond the distance threshold. Thus, in order to get good results, LSH
tends towards a brute force search over the candidate results. Despite
this, LSH is faster due to the efficiency of the hashing process.

%----------------------------------------------------------------------

\begin{figure}[t]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Skye dataset\label{skye-quality-mtree}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Kilmarnock dataset\label{kilmarnock-quality-mtree}}
\end{subfigure} \vspace{5mm}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
\vspace{-6mm}
\caption{F1 Measure on Skye dataset \\ with M-tree and all LSH configurations}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
\vspace{-6mm}
\caption{F1 Measure on Kilmarnock dataset \\ with M-tree and all LSH configurations}
\end{subfigure}
\caption{Linkage results on the demographic datasets.}
\label{demography-quality}
\end{figure}

%----------------------------------------------------------------------

\begin{table}[t]
\caption{Linkage quality on Skye dataset with distance threshold
$d= 2$.}
\label{comparison-of-results-demography-skye}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrr} \hline\noalign{\smallskip}
  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
%\parbox{0.2\textwidth}{\centering Linker} &
%\parbox{0.2\textwidth}{\centering Comparisons} &
%\parbox{0.17\textwidth}{\centering Precision} &
%\parbox{0.17\textwidth}{\centering Recall} &
%\parbox{0.17\textwidth}{\centering F1 Measure} \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2  & 3,109,250                   & 0.63      & 0.03   & 0.06       \\

LSH-5-2  & 10,412,496                  & 0.64      & 0.11   & 0.19       \\

LSH-10-2 & 53,874,127                  & 0.68      & 0.36   & 0.47       \\

LSH-5-5  & 36,566                     & 0.76      & 0.01   & 0.01       \\

LSH-10-5 & 129,873                    & 0.72      & 0.01   & 0.02       \\
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

%----------------------------------------------------------------------

\begin{table}[t]
\caption{Linkage quality on Kilmarnock dataset with distance
threshold $d= 2$.}
\label{comparison-of-results-demography-kili}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrrrr} \hline\noalign{\smallskip}
  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
%\parbox{0.2\textwidth}{\centering Linker} &
%\parbox{0.2\textwidth}{\centering Comparisons} &
%\parbox{0.17\textwidth}{\centering Precision} &
%\parbox{0.17\textwidth}{\centering Recall} &
%\parbox{0.17\textwidth}{\centering F1 Measure} \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2   & 99,145,887                    & 0.81      & 0.16   & 0.27       \\
LSH-5-2   & 130,721,338                   & 0.79      & 0.23   & 0.36       \\
LSH-10-2  & 177,168,848                   & 0.79      & 0.36   & 0.49       \\
LSH-5-5   & 239,368                      & 0.84      & 0.01   & 0.02       \\
LSH-10-5  & 855,431                      & 0.87      & 0.02   & 0.03 \\ 
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

%----------------------------------------------------------------------

\section{Conclusions and Future Work\label{sec-concl}}

In this paper we have demonstrated the efficacy of MSI in achieving
complete and efficient record linkage, without the need for complex
parameter tuning. In conclusion, this claim deserves some careful
unpacking. It is always possible to achieve high quality linkage using a
brute force approach. However the quadratic complexity of this approach
prevents its practical application for datasets of even moderate size.
We have shown that MSI techniques such as M-tree can deliver high
precision, high recall results that are the same as those delivered by
brute force. Furthermore this is achieved with fewer distance
comparisons, and consistently without the need for complex parameter
tuning.

We contrast this to traditional blocking and LSH-based approaches. Their
major drawback is that whilst they can produce extremely good results,
they can also produce extremely poor results. It was our observations of
low recall given by these approaches that originally led us to
experiment with M-trees.

We note that the good results yielded by both traditional blocking and
LSH are partly due to the fact that (in the limit) they tend towards
brute force as the number of records in the block increase. A second,
unexpected, result is that illustrated in
\Cref{comparison-of-results-cora}, namely that incomplete approaches
such as LSH can in some cases yield higher precision than that achieved
by a complete method such as M-tree. This is due to the incomplete
linker masking the inability of a classifier based solely on
record similarity to correctly classify high-similarity non-matches or
low-similarity matches.

The observation that LSH can yield results quickly at the cost of some
recall suggests that hybrid algorithms may be a potentially profitable
future direction for exploration. We hypothesise that it may be possible
to use an incomplete technique like LSH to form indices into a data
structure such as an M-tree. In this manner, M-tree nodes could be used
as the starting point for range-search queries and, in so doing, prune
the search space considerably to yield high quality results quickly.
We plan to develop such hybrid indexing approaches with the aim to
achieve fully automated, complete and efficient record linkage.

%----------------------------------------------------------------------

\bibliographystyle{splncs03}
\bibliography{paper.bib}

%----------------------------------------------------------------------

\end{document}

%----------------------------------------------------------------------

%text cut from earlier versions:

%The MIFile \cite{amato2014mi} approach is an inexact technique for performing similarity search based on metric spaces. Using  MI\_file,  each record is represented by the ordering of distances from a collection of reference objects.
% Peter: are these reference object also records? 
It is based on the idea that two records close to each other will have similar neighbours and will thus be represented by similar of distances
% Peter: unclear what you mean with: similar of distances
from the chosen reference objects. 
In order to create the indexes, for each value to be added to the
% peter: value -> record ?
data-structure, first the k nearest reference objects (governed by a fixed parameter) are found, each is given a score (drawn from 1,2,3... based on the reference object's position from the value).
% peter: this is unclear .. I think I understand k-NN: for each
% record you find the k nearest reference objects and you enumerate
% them
Next for each score, a  mapping is created in an inverted index mapping
% peter: unclear..? a mapping in an inverted index mapping?
from the value to the score of each reference object and a representation of the data associated with the reference object and the value inserted.
% peter: unclear: what is "the data" associated with the reference
% object.
In order to perform a \textit{nearest-neighbour} lookup, first the k nearest reference objects (bounded by a second fixed parameter).
% peter: this sentence reads incomplete..?
Next the objects closest to these reference objects are extracted from the inverted map by choosing the neighbours with the highest score.
% Peter: sorry, this is unclear as well. "the objects closest" ->
% records closest? And why highest scores..? Is the closest record
% to a reference object given the highest score (not the lowest
% score - i.e. not 1-closest, 2-closest etc)
Much of the search space may be eliminated by  keeping track of the scores that  can contribute to the nearest neighbours.
% peter: so I guess you use the triangular inequality and the reference
% objects to prune the search space.

\cite{Yu2016} surveys the use of string similarity in record linkage.
Algorithms used to compare two records are characterised as being
character-based or token-based. The first category, including approaches
such as Levenshtein~\cite{Levenshtein66}, uses character differences in
the strings being compared. The second, including techniques such as
Jaccard similarity, treats records as sets of tokens, and utilises set
similarity. There are also hybrids of the two approaches.
% Do we need the rest of this paragraph?
\cite{Yu2016} categorises linkage algorithms as being based on filtering,
verification, or thresholds.
% This is missing some context as to what is going on.
Filtering algorithms generate signatures for each record, construct
inverted lists containing records with common signatures, and prune
records with no common signatures.
% Distance of prefix to what? What is the data in this context?
Verification algorithms calculate the exact distances of a prefix of the
data, and use another technique to estimate the edit distance between
records. If the estimated distance is larger than the desired similarity
threshold, the candidate may be discarded.
% This seems a bit throw-away.
% Also, these aren't themselves linkage algorithms are they?
Threshold-based algorithms include M-trees, Trie-Join, Pass-Join and
Partenum.

%We will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
Sect.~\ref{sec-exp}.

% \begin{table}[t]
% \caption{Parameter settings used for the different datasets used in
%    the experiments,with $d$ the distance threshold used, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
%    shingle size, and the size and number of bands, respectively.}
%  \label{table-parameters}
%   \centering
%   \begin{scriptsize}
%   %\addtolength{\tabcolsep}{-0.5pt}
%   \begin{tabular}{ccccc}
%   \hline\noalign{\smallskip}
%   Dataset~ & $d$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ \\
%   name(s) & & & &  \\
%   \noalign{\smallskip} \hline \noalign{\smallskip}
%   CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$2$~ &
%     ~$[2,5,10]$~ & ~$[2,5,10]$~  \\
%   Isle of Skye &  \\
%   Kilmarnock  &  \\
%   \noalign{\smallskip} \hline
%   \end{tabular}
%   \end{scriptsize}
% \end{table}

% \textbf{Experimental Setup:}
% %
% We implemented all techniques described in Sect.~\ref{sec-approach}
% using Java (version 1.8.0\_144) and ran an extensive set of experiments on
% a compute server with 24 Intel(R) Xeon(R) processors running at 2.30GHz, 
% 128 GBytes of main memory and running Red Hat 4.8.5-11.