### 2018-02-14 paper.tex
After Final Revisions

\documentclass{llncs}

\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{url}
\usepackage{cleveref}
\usepackage{hhline}
\usepackage{subcaption}

\begin{document}

%----------------------------------------------------------------------

\title{Using Metric Space Indexing for Complete and Efficient Record Linkage}

%----------------------------------------------------------------------

\author{{\"O}zg{\"u}r Akg{\"u}n\inst{1} \and Alan Dearle\inst{1}
  \and Graham Kirby\inst{1} \and Peter Christen\inst{2}}

\institute{School of Computer Science, University of St Andrews,
St Andrews, Scotland. Contact: \email{\{ozgur.akgun, alan.dearle, graham.kirby\}@st-andrews.ac.uk}
\and
Research School of Computer Science, The Australian National University,\\
Canberra, Australia. Contact: \email{peter.christen@anu.edu.au}}

%----------------------------------------------------------------------

\maketitle

\begin{abstract}

Record linkage is the process of identifying records that refer to the
same real-world entities, in situations where entity identifiers are
unavailable. Records are linked on the basis of similarity between
common attributes, with every pair being classified as a link or
non-link depending on their similarity. Linkage is
usually performed in a three-step process: first, groups of similar
candidate records are identified using indexing, then pairs within the same
group are compared in more detail, and finally classified.
Even state-of-the-art indexing techniques, such as Locality Sensitive
Hashing, have potential drawbacks. They may fail to group together some
true matching records with high similarity, or they may group
records with low similarity, leading to high computational overhead.
We propose using \emph{metric space indexing} (MSI) to perform \emph{complete}
linkage, resulting in a parameter-free process
combining indexing, comparison and classification into a single step
delivering complete and efficient record linkage. An
evaluation on real-world data from several domains shows that
linkage using MSI can yield better quality than
current indexing techniques, with similar execution cost, without the
need for domain knowledge or trial and error to configure the process.

\end{abstract}

\keywords Entity resolution; data matching; similarity search;
         blocking.

%----------------------------------------------------------------------

\section{Introduction}
\label{sec-intro}

Record linkage, also known as entity resolution, data matching and
duplicate detection~\cite{Chr12}, is the process of identifying and
matching records that refer to the same real-world entities within or
across datasets. The entities to be linked are often people (such as
patients in hospital or customers in business datasets), but record
linkage can also be applied to link consumer products or bibliographic
records~\cite{Chr12}. Record linkage is commonly challenged by the lack
of unique entity identifiers (keys) in the datasets to be linked, which
prevents the use of a database join. Instead, the linkage of records
requires the comparison of the common attributes (or fields) that are
available within the datasets, for example the names, addresses and
dates of birth of individuals.

To overcome data quality issues such as typographical errors and
variations (common in name and address values~\cite{Chr12}),
approximate string comparison functions (e.g. edit distance, the
Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used to
compare record pairs, leading to a vector of similarities (one
similarity per attribute compared) for each pair. These
are used to classify the record pairs into
\emph{links} (where it is assumed both records correspond to
the same real-world entity) and \emph{non-links} (where they are
assumed to correspond to different entities). Various
classification methods have been employed in record
linkage~\cite{Chr12,Don15}, ranging from simple threshold-based to
sophisticated clustering, supervised classification techniques, and
active learning approaches~\cite{Wan15}.

Besides a lack of unique entity identifiers, and data
quality issues, linkage is also
challenged by dataset scale~\cite{Don15}. To avoid full pair-wise comparison of all possible
record pairs (quadratic in the dataset sizes),
blocking techniques, commonly known as \emph{indexing}~\cite{Chr12b},
are used. These split the datasets into smaller blocks in an efficient way, grouping together records that are likely to
correspond to the same entity. Only
records within blocks are then compared in detail.

While indexing allows efficient linkage of large
datasets~\cite{Don15}, scalability is at the cost of
reduced linkage quality, because potentially matching record pairs
are ignored, leading to lower
recall~\cite{Chr12}. Indexing techniques, discussed in more
detail later, range from simple
phonetic based blocking~\cite{Chr12} and sorting of the
datasets~\cite{Dra12} to locality sensitive hashing based
techniques~\cite{Kim10,Steorts2014}, and unsupervised~\cite{Kej13,Ram15} and supervised~\cite{Bil06,Mic06}
learning of optimal blocking schemes.

%----------------------------------------------------------------------

\begin{figure}[!t]
  \centering
  \includegraphics[width=1.0\textwidth]{figures/linkage-process}
  \caption{Overview of the steps of the traditional record linkage
           process (left side) and our proposed metric space
           indexing based approach (right side), as described in
           Sect.~\ref{sec-intro}, where records from two datasets,
           $\mathbf{D}_A$ and $\mathbf{D}_B$, are being linked.}
           \label{fig-rl-process}
\end{figure}

%----------------------------------------------------------------------

Traditional linkage systems that perform indexing prior to
comparison and classification (on the left in
\Cref{fig-rl-process}) add a further complexity. Indexing, comparison and classification are often conducted
using algorithms and parameters selected using
domain expertise, followed
by manual assessment of the linkage
outcomes~\cite{Chr12}. If the resulting link quality is too low for a certain application, the process is
repeated with different parameter settings or algorithms, giving a time-consuming iterative process. The choice of an appropriate indexing
technique and parameters (including which attributes to use in
indexing) will significantly affect the final linkage outcome.

We focus on approaches using a similarity threshold to
classify links. These are fundamentally limited by the extent
to which true matching records are similar, and true non-matches are
dissimilar---this is dataset-dependent. Within this domain, we define
a technique to be \emph{complete} if it guarantees to find all
record pairs within the specified threshold. Many
indexing techniques are incomplete, since they reduce computational
cost at the expense of potentially overlooking some true matches. By
definition, incomplete techniques yield lower recall than complete
ones. Conversely, and counter-intuitively, complete
techniques can yield lower precision with some datasets. This is
discussed further in Sect.~\ref{sec-approach}.

Metric space indexing (MSI) is a complete technique with
lower computational cost than a brute force approach. It allows
indexing, comparison and classification to be combined into a single
step (on the right in
\Cref{fig-rl-process}), making the process
simpler, more efficient and more effective than incomplete
approaches.

\textbf{Contribution:} Our primary contribution is the
novel application of MSI to achieve complete and efficient record
linkage, without the need for complex parameter tuning. We evaluate our
approach on several real-world datasets and
demonstrate its advantages over existing indexing techniques for
record linkage~\footnote{The motivation for this work is the Digitising Scotland
project~\cite{Dibben2012}, which aims to transcribe and
link all civil registration events recorded in Scotland between
1856 and 1973. This data set will include around 14
million birth records, 11 million death records and 4 million marriage
records.}.

\section{Related Work}
\label{sec-related}

We review relevant work in the areas of indexing for record
linkage (for recent surveys see~\cite{Chr12b,Pap16}), and metric space
indexing~\cite{Zezula2010}. Techniques to link records have been investigated for
over five decades~\cite{Fel69,New59}, with scalability
being an ongoing challenge as datasets grow in size and complexity.
Traditional blocking~\cite{Chr12b} uses a set of attributes (a
\emph{blocking key}) to insert records with the same value(s) in
their blocking key into the same block. Only records within the same
block are  compared to each other. To overcome variations and
misspellings, the values can be phonetically
encoded using functions such as Soundex, NYSIIS, or
Double-Metaphone~\cite{Chr12}. These convert a string into a code
according to its pronunciation, assigning the same code to similar
sounding names (such as `Gail' and `Gayle'). Multiple blocking keys
may also be used to deal with missing attribute values.

A different approach uses sorted neighbourhoods~\cite{Mon96}, where the datasets are sorted
according to a \emph{sorting key} (usually a concatenation of several attribute values). A sliding window is moved over the
datasets and only records within the window are compared. Techniques
that adaptively shrink or expand the window size based on the
characteristics of the sorting key values have been shown to improve
both linkage efficiency and quality~\cite{Dra12}.

These techniques are heuristics, requiring domain
knowledge, such as the choice of appropriate blocking or sorting keys.
Poor choices of blocking attributes result in records being inserted
into inappropriate blocks, and thus true matches being missed, giving \emph{incomplete} linkage.
Conversely, many pairs compared in a block may have low similarity, being non-matches, giving
\emph{inefficient} linkage.

Locality sensitive hashing (LSH), proposed for efficient
nearest-neighbour search in high-dimensional spaces~\cite{Ind98}, has
been used for record linkage indexing. Attribute
values are hashed multiple times, and blocks are created from those
records that share some hash values. \emph{HARRA}~\cite{Kim10} is a
linkage approach based on MinHash~\cite{Broder1997} and LSH
which blocks, compares, and then merges linked records iteratively. \cite{Steorts2014} evaluates two LSH variations, concluding
that to get good results, they must be tuned to the
datasets. This requires good ground
truth data which may be unavailable or expensive to obtain.

Metric space indexing (MSI) techniques~\cite{Zezula2010} support similarity search. 
They require
a distance measure between records, with certain properties including
the \emph{triangle inequality}~\cite{Zezula2010}. Similarity search
operations include \emph{range-search(\textbf{q},d)}, identifying all
records within a distance $d$
of a query record $\mathbf{q}$;
\emph{nearest-neighbour(\textbf{q})}, returning the record with smallest distance to $\mathbf{q}$; and
\emph{nearest-n(\textbf{q},n)}, returning the $n$ closest records to $\mathbf{q}$. Here we choose one MSI
structure, the M-tree~\cite{Ciaccia97indexingmetric}, and investigate
its efficacy for record linkage.
The M-tree is dynamically balanced. Every node contains
a reference to a record being indexed, a pointer to its parent, the
distance to its parent, and the node's radius. The radius of a node is
the distance from it to its furthest child. For a parent node with
radius \textit{r}, all its children may be visualised as being contained
within a ball of radius \textit{r} from it.

A linkage method using R-trees~\cite{Hjaltason1998}
was described in~\cite{Li2006}, demonstrating that high linkage
quality can be achieved using Jaccard similarity. \cite{Ciaccia97indexingmetric} shows that M-trees are
almost always more efficient than R-trees, hence their use here.

%----------------------------------------------------------------------

\section{Approach}
\label{sec-approach}

We address the following general linkage problem: for two
datasets $\mathbf{D}_A$ and $\mathbf{D}_B$, we wish to find, for each
record in $\mathbf{D}_A$, all the records in $\mathbf{D}_B$ that match
it with regard to a certain distance threshold $d$ (i.e.\ have a
distance of $d$ or less). We compare several linkage algorithms: traditional blocking, an incomplete similarity search method,
LSH-MinHash, and a complete method, M-tree. We also use a
complete brute force technique as a baseline, though
this can only feasibly be applied to our smallest dataset. All
experiments have a number of parameters to configure the search space
and algorithm behaviour, including the distance function and the threshold, $d$,
specifying the maximum distance for
two records to be classified as a link (i.e.\ referring to the same
entity).
We focus on a single distance function in these experiments, to constrain the experimental space. In
Sect.~\ref{sec-concl} we return to the selection of alternative distance functions.

\textbf{Brute force:}
Every record in
$\mathbf{D}_A$ is compared with every record in $\mathbf{D}_B$. Each pair is
classified as a link if the distance between the records is less than or equal to
the threshold. This always finds
all links, with complexity $ O(|\mathbf{D}_A| \times |\mathbf{D}_B|)$.

\textbf{Traditional blocking:}
The parameters are the set of blocking keys and
(optionally) the phonetic encodings applied to each attribute. These are
selected as described in~\cite{Chr12b}, exploiting knowledge of the
domain and of the data, and chosen with the intention of
giving the best possible results. Each record in
$\mathbf{D}_A$ is placed into the appropriate block based on its
blocking key value. The algorithm then iterates over the records in
$\mathbf{D}_B$, and for each one, compares it with each of the records
from $\mathbf{D}_A$ in the block with the same blocking key value.

\textbf{LSH-MinHash:}
The parameters for LSH-Minhash are~\cite{Broder1997} \emph{shingle
size} ($l_{ss}$), \emph{band size} ($l_{bs}$) and \emph{number of
bands} ($l_{nb}$). First, the attributes of each record in $\mathbf{D}_A$
are concatenated, and the result \emph{shingled} into a set of
n-grams with $n = l_{ss}$. Next, a set of deterministically generated
hash functions are applied to each n-gram in the set and the smallest
result (the MinHash) of each hash application is added to a signature
for the record. The number of hashes used, and thus the size of the
signature, is set to $l_{nb} \times l_{bs}$. Finally, the signature is
split into $l_{nb}$ bands and the values from each band are hashed
again to create a number of keys. The original record is added to a
map associated with each of the keys. 
To perform linkage, the algorithm iterates over the records in
$\mathbf{D}_B$. Each record is hashed as described
above, to obtain a set of keys. Each key is
looked up in the data structure, and the associated records from
$\mathbf{D}_A$ added to the result set. Finally,
the record from $\mathbf{D}_B$ is compared in turn with each record
in the result set, with the pair being classified as a link
or non-link based on their distance.

In some circumstances, incomplete approaches such as traditional
blocking and LSH-MinHash can yield higher precision than complete
techniques. This can occur when a significant
number of non-matches nonetheless have high similarity. In this
situation, the fact that an incomplete technique omits consideration
of some potential links can serve to improve precision, since a
classification decision based on a certain similarity threshold is
incorrect for high-similarity non-matches. By definition, recall can
never be higher for incomplete techniques.

\textbf{M-tree:} The linkage
algorithm has no additional parameters. As with
\emph{LSH-MinHash}, each record in $\mathbf{D}_A$ is inserted
into an M-tree. To perform linkage, the algorithm iterates over each
record $\mathbf{b} \in \mathbf{D}_B$. A
\emph{range-search(\textbf{b},d)} operation is performed on the M-tree,
passing the distance threshold $d$ as the second parameter. All the
returned records are directly classified as links.

%----------------------------------------------------------------------

\section{Experiments and Results}
\label{sec-exp}

We now describe the datasets and method used in our evaluation~\footnote{The raw data, additional figures and the source
code required to run these experiments can be downloaded from:
(link removed for double blind review).}.
\label{sec-data}
We used three datasets from two domains in our experiments, as
summarised in Table~\ref{table-datasets}. The first is
\emph{Cora}~\cite{Cora2017}, which contains 1,295 records that refer
to 112 machine learning publications. Cora is commonly used as a
benchmark dataset in the literature for assessing linkage algorithms.
Ground truth is provided via a unique \emph{paper\_id} identifier of
the form ``blum1993". In this experiment linkage is performed over
the same set of records (i.e.\ a de-duplication~\cite{Chr12}).

%----------------------------------------------------------------------

\begin{table}[t]
\caption{Characteristics of datasets used in the experiments.}
 \label{table-datasets}
  \centering
  \begin{scriptsize}
  \begin{tabular}{ccccc}
  \hline\noalign{\smallskip}
  Dataset~ & ~Records in~& ~Records in~ & ~Number of true~& ~Entities \\
    name(s)  & ~dataset $\mathbf{D}_A$~ & ~dataset $\mathbf{D}_B$~ &
    matching pairs & linked \\
  \noalign{\smallskip} \hline \noalign{\smallskip}
  Cora & 1,295 & 1,295 & 17,184 & Publication--Publication\\
  ~Isle of Skye~ & 17,612 & 12,284& 2,900 & Birth--Death \\
  Kilmarnock & ~38,430~ & ~23,714~ & ~8,300~ & ~Birth--Death~ \\
  \noalign{\smallskip} \hline
  \end{tabular}
  \end{scriptsize}
\end{table}

%----------------------------------------------------------------------

The other two datasets are historical Scottish records of vital events
(birth, marriages and deaths), one registered on the \emph{Isle of
Skye}, a rural district, and the other records from
\emph{Kilmarnock}, an industrial town. These datasets were created, curated and linked by
historical demographers~\cite{reid2006,reid2002}. Both include the names and
genders of individuals and their parents. Ground truth was
generated by the demographers based on their extensive domain
knowledge.

In all of our experiments we use a single distance metric: the
sum of the attribute-level Levenshtein~\cite{Levenshtein66} edit distances.

\subsection{Cora Results}

We perform linkage on the Cora dataset using all
approaches presented in this paper: brute force, traditional blocking,
LSH and M-tree, using several selected configurations for blocking
and LSH. The distance threshold is varied between 0 and 250~\footnote{Relatively high Levenshtein edit distances are
included since Cora includes a number of low-similarity true
matches.}. For traditional blocking, the following attributes are
used individually as blocking keys: \emph{author}, \emph{title},
\emph{venue}, \emph{location}, \emph{publisher} and \emph{year}. We
also use a combined blocking key comprising all attributes.

\Cref{cora-quality} shows the precision, recall, and
F-measure~\cite{Chr12} for various thresholds. As expected, low
thresholds give high precision and low recall, and the reverse for high
thresholds. Brute force and M-tree give identical results, as
expected. The best linkage quality, with an F-measure of around 0.7, is
achieved by several linkers, including brute force, M-tree, blocking
on \emph{authors}, blocking on all attributes, and two of the LSH
configurations. All of these give similar overall results, apart from
blocking on \emph{authors}, which gives much better quality at very high
distance thresholds. This is due to the incomplete nature of the
approach, avoiding comparisons of significant numbers of high-similarity
non-matches and thus avoiding these becoming false positives and keeping
precision high.

For a more detailed investigation of selected linkers, the brute force
approach is used to establish a good threshold value for the Cora
dataset. The maximum F-measure is observed at a threshold value of
$d=70$. This value is dataset-dependent; for different datasets the
maximum F-measure will occur at different thresholds. In the rest of
this section we fix the threshold value at $d=70$.

\Cref{comparison-of-results-cora} shows greater detail for selected
linkers, showing the parameters for the experiment, the number of
distance comparisons made, and the precision, recall and F-measure
achieved by each algorithm. In the \emph{Linker} column the algorithm
name is followed by its parameters: for LSH the number of the bands
followed by the band size, and for traditional blocking the attributes
used for blocking. The number of distance comparisons is reported as a
machine-independent proxy for execution cost, since code profiling
shows that distance calculations are dominant.

M-tree yields the same linkage quality as brute force, although using a
significantly lower number of comparisons. This is as expected, since
both techniques are complete. Several of the incomplete linkers give
similar quality, for example \emph{LSH-2-2}, \emph{LSH-5-2},
\emph{LSH-10-2} and \emph{Block-combined}. These, and a number of other
incomplete linkers, give better precision than the complete techniques.
This is due to high-similarity non-matches, as discussed in
Sect.~\ref{sec-approach}. Although several of the incomplete linkers
give as good quality as M-tree, and in some cases at lower cost, this is
offset by the need to select appropriate configuration parameters. Some
other linkers give very poor results.

%----------------------------------------------------------------------

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
\vspace{-6mm}
\caption{Brute Force}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
\vspace{-6mm}
\caption{M-tree}
\end{subfigure} \vspace{5mm}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
\vspace{-6mm}
\caption{Blocking on `authors' attribute}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
\vspace{-6mm}
\caption{Blocking on `title' attribute}
\end{subfigure} \vspace{5mm}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
\vspace{-6mm}
\caption{Blocking on `publication year' attr.}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
\vspace{-6mm}
\caption{Blocking using the union of all attrs.}
\end{subfigure} \vspace{5mm}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-2}
\vspace{-6mm}
\caption{LSH using 2 bands of size 2}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-10}
\vspace{-6mm}
\caption{LSH using 2 bands of size 10}
\end{subfigure} \vspace{5mm}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-2}
\vspace{-6mm}
\caption{LSH using 10 bands of size 2}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-10}
\vspace{-6mm}
\caption{LSH using 10 bands of size 10}
\end{subfigure}

\caption{Linkage results on the Cora dataset.}
\label{cora-quality}
\end{figure}

%----------------------------------------------------------------------

\begin{table}[ht]
\caption{Linkage quality on Cora dataset with distance threshold
  $d= 70$.}
  \label{comparison-of-results-cora}
  \centering
  \begin{footnotesize}
  \begin{tabular}{lrrrr} \hline\noalign{\smallskip}
  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
  \noalign{\smallskip} \hline \noalign{\smallskip}
Brute Force        & 1,677,025                 & 0.84      & 0.57   & 0.68  \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree              &  902,693                  & 0.84      & 0.57   & 0.68       \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2            &  192,199	& 0.95	& 0.47	& 0.63       \\
LSH-5-2            &  342,849	& 0.91	& 0.55	& 0.69       \\
LSH-10-2           &  513,947	& 0.88	& 0.57	& 0.69       \\
LSH-2-5            &  14,329	& 0.99	& 0.28	& 0.43      \\
LSH-5-5            &  22,057	& 0.99	& 0.36	& 0.53       \\
LSH-10-5           &  26,167	& 0.98	& 0.4	& 0.57      \\
LSH-2-10           &  4,711	& 1	& 0.15	& 0.27    \\
LSH-5-10           &  6,501	& 1	& 0.19	& 0.32      \\
LSH-10-10          & 10,627  	& 0.99	& 0.27& 	0.43       \\
\noalign{\smallskip} \hline \noalign{\smallskip}
Block-year         &  115,893                  & 0.99      & 0.35   & 0.51       \\
Block-authors      &   11,039                   & 0.94      & 0.16   & 0.28       \\
Block-title        &   27,407                   & 0.95      & 0.42   & 0.58       \\
Block-venue        &   36,647                   & 0.85      & 0.29   & 0.44       \\
Block-location     & 1,009,957                 & 0.83      & 0.43   & 0.57       \\
Block-publisher    &  833,079                  & 0.85      & 0.44   & 0.58       \\
Block-combined     & 1,214,269                 & 0.84      & 0.56   & 0.67       \\
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

%----------------------------------------------------------------------

\subsection{Demographic Dataset Results}

Birth records were linked to death records, separately for the Skye and
Kilmarnock datasets, using M-tree and a range of LSH configurations. It
was not computationally feasible to run the brute force linker. Of the
incomplete linkers, LSH was selected as it gave slightly better results
for Cora. The shingle size was set to $l_{ss}=2$ for all the LSH
experiments reported, as this was found to give good results and LSH
was not especially sensitive to this parameter. Results for other
shingle sizes are omitted from this paper for brevity. A lower range
of distance thresholds was explored, based on domain knowledge of the
datasets.

\Cref{demography-quality} plots (a) and (b) show the M-tree precision,
recall, and F-measure for various thresholds. In both datasets, the best
F-measure values are obtained with a low distance threshold of $d=2$.
Plots (c) and (d) compare the F-measure curves for M-tree with those
obtained from a range of LSH configurations. The best F-measure value
for M-tree is higher than that of any of the LSH configurations, for
both datasets. This demonstrates both the competitiveness of M-tree
with respect to linkage quality, and its important characteristic of
being parameter-free---the linkage quality is obtained without the
need to tune for the dataset.

Tables~\ref{comparison-of-results-demography-skye}
and~\ref{comparison-of-results-demography-kili} show greater detail for
selected linkers. In both cases the F-measure achieved is better for
M-tree than any of the LSH linkers. The better linkage quality achieved
by M-tree is largely due to recall for M-tree being much higher than for
any of the LSH configurations. In most cases, LSH out-performs M-tree in
terms of precision. More significantly, LSH linkage quality is heavily
dependent on the configuration parameters. For plausible settings for
the number of bands and band size, F-measure varies from $0.01$
(extremely poor) to $0.47$ (relatively good) for Skye and from $0.03$
to $0.49$ for Kilmarnock. In both cases \emph{LSH-10-2} performs best,
but since this is data-dependent there is no guarantee that these
parameters would work well with another dataset.

The number of distance comparisons varies dramatically among the various
linkers. M-tree always performs the most comparisons, since they are
intrinsic to the \emph{range-search} algorithm. The core part of the LSH
linker performs Jaccard similarity comparisons and hashing; distance
comparisons are only performed in the final step to determine whether a
candidate pair is a link. The LSH configurations yielding the best
results perform distance comparisons of the same order of magnitude as
M-tree. This indicates that the good LSH linkers return many candidates
beyond the distance threshold. Thus, in order to get good results, LSH
tends towards a brute force search over the candidate results. Despite
this, LSH is faster due to the efficiency of the hashing process.

%----------------------------------------------------------------------

\begin{figure}[t]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Skye dataset\label{skye-quality-mtree}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Kilmarnock dataset\label{kilmarnock-quality-mtree}}
\end{subfigure} \vspace{5mm}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
\vspace{-6mm}
\caption{F1 Measure on Skye dataset \\ with M-tree and all LSH configurations}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
\vspace{-6mm}
\caption{F1 Measure on Kilmarnock dataset \\ with M-tree and all LSH configurations}
\end{subfigure}
\caption{Linkage results on the demographic datasets.}
\label{demography-quality}
\end{figure}

%----------------------------------------------------------------------

\begin{table}[t]
\caption{Linkage quality on Skye dataset with distance threshold
$d= 2$.}
\label{comparison-of-results-demography-skye}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrr} \hline\noalign{\smallskip}
  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2  & 3,109,250                   & 0.63      & 0.03   & 0.06       \\

LSH-5-2  & 10,412,496                  & 0.64      & 0.11   & 0.19       \\

LSH-10-2 & 53,874,127                  & 0.68      & 0.36   & 0.47       \\

LSH-5-5  & 36,566                     & 0.76      & 0.01   & 0.01       \\

LSH-10-5 & 129,873                    & 0.72      & 0.01   & 0.02       \\
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

%----------------------------------------------------------------------

\begin{table}[t]
\caption{Linkage quality on Kilmarnock dataset with distance
threshold $d= 2$.}
\label{comparison-of-results-demography-kili}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrrrr} \hline\noalign{\smallskip}
  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2   & 99,145,887                    & 0.81      & 0.16   & 0.27       \\
LSH-5-2   & 130,721,338                   & 0.79      & 0.23   & 0.36       \\
LSH-10-2  & 177,168,848                   & 0.79      & 0.36   & 0.49       \\
LSH-5-5   & 239,368                      & 0.84      & 0.01   & 0.02       \\
LSH-10-5  & 855,431                      & 0.87      & 0.02   & 0.03 \\ 
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

%----------------------------------------------------------------------

\section{Conclusions and Future Work\label{sec-concl}}

In this paper we have demonstrated the efficacy of MSI in achieving
complete and efficient record linkage, without the need for complex
parameter tuning. In conclusion, this claim deserves some careful
unpacking. It is always possible to achieve high quality linkage using a
brute force approach. However the quadratic complexity of this approach
prevents its practical application for datasets of even moderate size.
We have shown that MSI techniques such as M-tree can deliver high
precision, high recall results that are the same as those delivered by
brute force. Furthermore this is achieved with fewer distance
comparisons, and consistently without the need for complex parameter
tuning.

We contrast this to traditional blocking and LSH-based approaches. Their
major drawback is that whilst they can produce extremely good results,
they can also produce extremely poor results. It was our observations of
low recall given by these approaches that originally led us to
experiment with M-trees.

We note that the good results yielded by both traditional blocking and
LSH are partly due to the fact that (in the limit) they tend towards
brute force as the number of records in the block increase. A second,
unexpected, result is that illustrated in
\Cref{comparison-of-results-cora}, namely that incomplete approaches
such as LSH can in some cases yield higher precision than that achieved
by a complete method such as M-tree. This is due to the incomplete
linker masking the inability of a classifier based solely on
record similarity to correctly classify high-similarity non-matches or
low-similarity matches.

We have focused on a single distance function: the sum of attribute-level Levenshtein distances. This gives a straightforward intuition of record-level distances, but it is more common to normalise metrics to the range 0-1.

Many distance functions, such as Jaro-Winkler, are not metric, and therefore cannot be used with MSI techniques. Care must be taken to preserve metric properties when combining metrics over individual fields, as is highlighted in~\cite{Yujian2007}, since this may yield a function that is not metric. It is then possible for MSI techniques to yield results that are subtly incorrect.

Different metrics can give different distributions of inter-record distances, which can affect both the linkage results and the number of comparisons made, and hence the efficiency of the algorithm.
Datasets with low variation in inter-record distances are said to have high \textit{intrinsic dimensionality}, and tend to require high numbers of comparisons.
The Scottish vital event datasets, combined with Levenshtein-based metrics, appear to have high intrinsic dimensionality.
We are now, therefore, investigating the application of various different metrics to this domain, including Jensen-Shannon, Cosine and Structured Entropic Distance, as described in~\cite{Connor2016}. We are also investigating the applicability of a novel technique for dimensionality reduction~\cite{Connor2017}.

%----------------------------------------------------------------------

\section{Acknowledgements}

This work was supported by ESRC grants ES/K00574X/2 ``Digitising Scotland''
and ES/L007487/1 ``Administrative Data Research Centre---Scotland''.

We thank Alice Reid of the University of
Cambridge and her colleagues, especially Ros Davies and Eilidh Garrett,
for the work undertaken on the Kilmarnock and Isle of Skye databases.

%----------------------------------------------------------------------

\bibliographystyle{splncs03}
\bibliography{paper.bib}

%----------------------------------------------------------------------

\end{document}

%----------------------------------------------------------------------